---
title: 'AnSeq_groupProject'
output: pdf_document
---

-   [Introduction]

    -   [Data]

-   [Exploratory Data Analysis (EDA)](#part-1---exploratory-data-analysis-eda)

    -   [Visualization]
    -   [Discussion](#discussion-2)

-   [Indicators]

-   [Simple Model]

    -   [Discussion](#discussion-2)
    -   [Fitting](#fitting-1)

-   [Exponential Smoothing]

    -   [Discussion](#discussion-2)
    -   [Fitting](#fitting-1)
    -   [Analysis](#analysis-1)
    -   [Comparison with simple model]

-   [ETS AND AUTO-ARIMA]

    -   [Fitting](#fitting-1)
    -   [Analysis](#analysis-1)
    -   [Discussion](#discussion-2)
    -   [Comparison with other models]

-   [Conclusion]

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(message = FALSE)

#needed libraries
library(Mcomp)
library(ggplot2)
```

# Introduction (Kevin)

The goal of this project is to sharpen our time series forcasting skills as well as the interpretation of the generated predictions.

## Data

We use the data called n1490 which are datapoints for amazon shipments between the years 1990 to 1996

```{r}
subset(M3, "monthly")["N1483"] %>% str()

data <- subset(M3, "monthly")["N1483"]

autoplot(data$N1483)

```

# Exploratory Data Analysis (EDA) {#part-1---exploratory-data-analysis-eda} (This)

## Visualization

```{r}

```

## Discussion

# Indicators (Kevin)

RMSE

# Simple Model (This)

## Discussion

blabla

## Fitting

```{r}

```

# Exponential Smoothing (Kevin)

## Discussion {#discussion-2}

blabla

- problem of linear approach: prediciting far into future will get high uncertainty
- damped will end converge


SHOW RESIDUALS
## Fitting {#fitting-1}

```{r}




```

## Analysis

```{r}

```

## Comparison with simple model

blabla

```{r}

```

# ETS AND AUTO-ARIMA (Kevin)

## Fitting

```{r}

```

## Analysis {#analysis-1}

```{r}

```

## Discussion

blabla

## Comparison with other models

```{r}

```

# Conclusion (This)

blabla
