---
title: 'AnSeq_groupProject'
output: pdf_document
---

-   [Introduction]

    -   [Data]

-   [Exploratory Data Analysis (EDA)](#part-1---exploratory-data-analysis-eda)

    -   [Visualization]
    -   [Discussion](#discussion-2)

-   [Indicators]

-   [Simple Model]

    -   [Discussion](#discussion-2)
    -   [Fitting](#fitting-1)

-   [Exponential Smoothing]

    -   [Discussion](#discussion-2)
    -   [Fitting](#fitting-1)
    -   [Analysis](#analysis-1)
    -   [Comparison with simple model]

-   [ETS AND AUTO-ARIMA]

    -   [Fitting](#fitting-1)
    -   [Analysis](#analysis-1)
    -   [Discussion](#discussion-2)
    -   [Comparison with other models]

-   [Conclusion]

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(message = FALSE)

#needed libraries
library(Mcomp)
library(ggplot2)
```

# Introduction (Kevin)

The goal of this project is to sharpen our time series forcasting skills as well as the interpretation of the generated predictions.

## Data

We use the data called n1490 which are datapoints for amazon shipments between the years 1990 to 1996

```{r}
subset(M3, "monthly")["N1483"] %>% str()

data <- subset(M3, "monthly")["N1483"]

autoplot(data$N1483)

```

# Exploratory Data Analysis (EDA) {#part-1---exploratory-data-analysis-eda} (This)

## Visualization

```{r}
ggsubseriesplot(data$N1483$x) +
  ggtitle("Seasonal plot: Monthly shipments")

ggseasonplot(data$N1483$x, year.labels=TRUE, year.labels.left=TRUE) +
  ggtitle("Seasonal plot: Monthly shipments")

ggAcf(data$N1483$x) 

gglagplot(data$N1483$x)

```

## Discussion

The following section outlines the selected dataset in an explanatory manner to provide a better understanding of the underlying data and to discover patterns and anomalies in the chosen time series.

The chosen dataset consists of 126 entries containing shipment data from 1985 to 1995 with a monthly frequency. As being part of the M3 competition data set, the given time series is already split up in a training and testing set containing 108 and 18 entries respectively.

As represented in the chart N1483, the chosen time series indicates a strong trend as the "amount" of shipments are growing constantly from 1985 to 1995. Furthermore, the graph implies that the time series has no seasonality. The mentioned constant trend and lack of seasonality can also be observed in appendix x and y. Appendix x clearly depicts that there is a constant growing trend in each month over the years in the given time series. In addition, this constant increasing trend over the months underlines that there is no seasonality in sequential data. The same observation can be made in appendix y, which clearly shows that there is now no seasonality as the plotted yearly shipments have no common peaks or lows. Instead the yearly lines tend to increase during the year. This growth over the year leads to the gaps between the line graphs, which represent the shipment over a year. Those described features emphasize the fact that the time series has an increasing trend and no seasonality.

Cycling trend ? -- Meiner Meinung nach nicht

The correlogram as well as the seasonal lag plot in appendix x and z, strengthen the assumption of an increasing trend and a missing seasonality. The autocorrelogram (ACF) in appendix x manifest that the autocorrelations for smaller lags are large and positive, which is typically for trendy data. ACF decreases over time - Rückschlüsse besprechen mit Kevin

GGlagplot mit kevin besprechen alle nahe zusammen jedoch eine Verschiebung zu sehen

# Indicators (Kevin)

RMSE

# Simple Model (This)

## Discussion

blabla

## Fitting

```{r}

```

# Exponential Smoothing (Kevin)

## Discussion {#discussion-2}

blabla

-   problem of linear approach: prediciting far into future will get high uncertainty
-   damped will end converge

SHOW RESIDUALS \## Fitting {#fitting-1}

```{r}




```

## Analysis

```{r}

```

## Comparison with simple model

blabla

```{r}

```

# ETS AND AUTO-ARIMA (Kevin)

## Fitting {#fitting-1}

```{r}

```

## Analysis {#analysis-1}

```{r}

```

## Discussion

blabla

## Comparison with other models

```{r}

```

# Conclusion (This)

blabla
